%************************************************
\chapter{Experiment Data}
\label{chapter:experiment_data}
%************************************************

This appendix chapter includes empirical data gathered using the
reflective tracing features of the underlying Funk garbage collected
memory system, upon which the virtual operating system is built.  This
low level memory tracing functionality allows easily tracing events at
any level of abstraction in the entire system using the Funk causal
reflective tracing features.  \autoref{table:bootup} shows a list of
figures showing data from an experiment where the AI simply booting up
and then perceiving the world over time.  This experiment is a sort of
control reference that shows that the AI memory usage and execution is
stable if it does not have any goals and the physical world is not
changing.  \autoref{table:reflective_learning} shows a list of figures
showing data from an experiment with the AI reflectively tracing, and
building models of the deliberative planning process.
\autoref{table:no_reflective_learning} shows a list of figures showing
data from an experiment with the reflective learning disabled.
Comparing the execution of the experiment with no reflective learning
with that of reflective learning shows a rough estimate of how much
more memory is used when the reflective learning about the
deliberative planning machine is enabled.

\experimentdatatable[bootup]{an experiment that simply boots up the AI
  and simply allows it to perceive the physical world and not
  stimulating it to achieve any goals} {This experiment serves as a
  control for the subsequent two experiments.  This experiment shows
  that when the AI is not given any deliberaive goals, it simply
  learns from its initial perceptions, and when these do not change,
  it has nothing to learn and memory usage is stable.}

\experimentdatatable[reflective_learning]{an experiment with the AI
  reflectively tracing, and building models of the deliberative
  planning process} {This experiment serves as a demonstration of the
  complete AI system running, learning from its physical perceptions,
  trying to accomplish the given deliberative goal of stacking two
  blocks, and also reflectively perceiving and learning to predict the
  effects of its deliberative actions, including how planning can lead
  to execution failures.}

\experimentdatatable[no_reflective_learning]{an experiment without the
  AI reflectively tracing, and building models of the deliberative
  planning process} {This experiment serves to show the memory
  performance of the system when the reflective perception, learning,
  and prediction of deliberative actions is disabled.}


\clearpage
\section{Figures from Mind Bootup Experiment}

%\experimentdatafigures[70]{bootup}{an experiment that simply boots up
%  the AI, allowing it to contine to perceive the physical world over
%  time without stimulating it to achieve any goals.  This experiment
%  shows that the AI learns from its initial perceptions, and when
%  these do not change, it has nothing to learn and memory usage is
%  stable.}

\clearpage
\section{Figures from Reflective Learning Experiment}

%\experimentdatafigures[140]{reflective_learning}{an experiment with
%  the AI reflectively tracing, and building models of the deliberative
%  planning process}

\clearpage
\section{Figures from Learning Experiment without Reflection}

%\experimentdatafigures[90]{no_reflective_learning}{an experiment
%  without the AI reflectively tracing, and building models of the
%  deliberative planning process}

