%************************************************
\chapter{From Simulation to Implementation}
\label{chapter:from_simulation_to_implementation}
%************************************************

\section{Digital Abstraction}

Computer simulations are based on the \emph{digital abstraction}, an
assumption that provides the most basic symbols of a computational
model, ``$1$'' and ``$0$''.  I will refer to the capacity to arrange
symbols in Space as \emph{computer memory}.  I will refer to a
specific arrangement of symbols in computer memory as \emph{data}.
While computer memory is often thought of as arising from physical
objects, such as transistors, the specific physical derivation of
these symbols is not important, given the modelling assumption of
digital abstraction.  For example, computer memory can be written down
on paper and manipulated by hand, once the idea is understood, it is a
tool for thinking, separately from any physical implementation.  This
assumption is necessary when using computers in modelling work.
Computer memory is an artificial static arrangement of symbols in
Space.

\section{The Combinational Device}

The digital abstraction is not all that is assumed when using a
computer to simulate a model.  The digital abstraction provides the
static symbols in Space, so let us now focus on the dynamic qualities
of Duration.  In simulating a model with a computer, there is a
predetermined discrete transition from the past to the future.  In the
past, there is an arrangement of symbols in Space and there is a
predetermined activity in Duration that causes these symbols in Space
to be arranged in a different specific configuration in the future.
This discrete activity is designed to be exactly the same for every
transition from the past to the future.  This logical predetermined
transition is referred to as \emph{the combinational device}.  The
combinational device maps any specific combination of symbols
Spatially arranged in the past to a specific Spatial arrangement in
the future.

\section{Digital Reflection}
\label{section:digital_reflection}

Because the combinational device is always understood to be the exact
same activity, there is not a non-tautological reason to symbolically
reference this activity in the computer memory.  Because this
transitional activity is always understood to be exactly the same, the
symbolic cause that refers to the activity during the transition from
the past to the future would be predefined and thus always the same,
making no distinctions.  I will refer to the ability to symbolize the
ongoing dynamic quality of the digital transition from the past to the
future as \emph{digital reflection}.  As I've described, it is a
meaningless and tautological pursuit for computers to exhibit digital
reflection.  Digital reflection is one of two forms of actual
reflection that are available to a computer.

\section{The Programmer}

Symbols have meaning only in reference to the ongoing qualities of
Duration.  So, then, the question becomes: if digital reflection is a
meaningless tautological pursuit, to what qualities of Duration do
digital symbols meaningfully and non-tautologically refer?  The
symbols in computer memory are programmed by a programmer that uses
these symbols as references to the qualities of Duration that are
actively ongoing and currently being experienced.  The computer is an
artificial tool of thought for the programmer, who provides the
meaningful reference to reality for the symbols in the computer
memory.  Again, a computer does not create symbols that reference
reality; the programmer is the reference to reality, while the
computer is an artificial tool of thought.
