%************************************************
\chapter{Introducing Simulation}
\label{chapter:introducting_simulation}
%************************************************

Keeping distinctions between the continuous dynamic referent and
discrete static symbols, a model can be constructed which allows $n$
layers of reflection to be described.  Non-reflective AI models learn
by correcting one immediate cause of a failure; however, given a
layered model, New opportunities for considering one failure to also
be failures in the reflective layers ``above'' and ``below'' the
original failure can be described, resulting in an arbitrary number of
reflective learning opportunities from a single failure.

I will not attempt to give a complete representation for a general
problem domain nor a complete description of a general problem solver.
My work here leaves the unmodellable as known to be just that.
Thinking as a dynamic activity uses models as useful tools with the
clear understanding that these models are static.  Further, and
critically important for the advancement of the field of AI, when
describing reflective thinking, it is imperative to have an explicit
awareness of thinking being limited to manipulating the static.  I
will therefore refer to an AI model as simply an ``AI'' with the
shared understanding that I am referring to a static model.

The simulation model is an extension of the model of mind that
includes a mathematical description of a discrete ``state'' of the
model in a discrete time.  Neither the model of mind nor the
simulation model changes over time.  The simulation model therefore
makes reference to a discretely stepped state that is used to simulate
the necessarily separate, static model of mind.  The term
``simulation'' is thus used as a general reference to the dynamic
activity in Duration that manipulates and steps the state of the
simulation model.  In describing the simulation model, it is important
to not confuse the dynamic activity of simulation with the model or
the state of the model, which are both static at any given point in
time.  The term ``simulation'' is the only reference to activities in
Duration.  This distinction is necessary for the simulation model to
not be limited to simulating a specific kind of activity.

For example, consider a simulation of topological proof.  A
mathematician can ``simulate'' the rules of topological proof by first
seeing a static arrangement of symbols on paper.  These symbols can be
manipulated in the process of proving or disproving the initial
topological statement.  There is no existing logical or mathematical
formulation of the activities of topological proof.  I describe how a
symbolic state space can be added to the model of mind for the
purposes of simulation.  How exactly this simulation is done is left
until the next part, where the activity of simulation is assumed to be
computational.

The activities in Duration in the model of mind exist prior to their
symbolization, and since it is obviously not possible to refer to
something prior to symbolizing it, the assumption that these
activities have already been symbolized is necessary in order to
mathematically describe the state of the simulation model.  Defining
the simulation model requires a description in terms of symbols for
the ``undescribed activities in Duration.''  Each assumption made in
symbolizing the ongoing activities in Duration restricts the
simulation model from being a model of those assumptions because only
the symbolic result of those assumptions is available for the
simulation of the activity of reflection.

