%************************************************
\chapter{The Dynamic and the Static}
\label{chapter:the_dynamic_and_the_static}
%************************************************

In this dissertation I focus on a description of thinking that
reflectively learns to accomplish goals.  I begin by describing a
clear distinction between the dynamic and the static, which is
\emph{the} fundamental key distinction in my model.  My model of
reflective thinking grows quickly within this initial foundational
conception, which, as it begins reflectively, leads necessarily and
naturally to the goals of this dissertation.  This first part of the
dissertation gives a non-technical, but complete, understanding of my
reflective model of mind.

I will not attempt to give a complete representation for a general
problem domain nor a complete description of a general problem solver.
My work here leaves the unmodellable as known to be just that.
Thinking uses models as useful tools with the clear understanding that
these models are static.  Further, and critically important for the
advancement of the field of AI, when describing reflective thinking,
it is imperative to have an explicit awareness of thinking being
limited to manipulating the static.  I will therefore refer to an AI
model as simply an ``AI'' with the shared understanding that I am
referring to a static model.

Non-reflective AI models learn by correcting one immediate cause of a
failure; however, given a layered model, I describe new opportunities
for considering one failure to also be failures in the reflective
layers ``above'' and ``below'' the original failure, resulting in an
arbitrary number of reflective learning opportunities from a single
failure.

In \autoref{part:the_implementation}, I will describe an example of a
computational implementation of reflective thinking.  The
implementation is a layered cognitive architecture with three working
layers.  I will focus on a very simple physical block stacking problem
domain as a tool for explaining simply, through examples, the layered
classes of reflective learning.  In describing the working model
implementation, I will maintain an awareness of the explicit
distinction between a static model and the dynamic referent.  I have
categorized my focus in this way only as an example of the
computational description.

\section{Dynamically Grounded Description}

A grounded description references the dynamic.  Words and language
have grounding only in this way.  The further ability of words to also
reference other words allows the additional possibility of creating a
referential circularity in language.  An illusion of a primitive
grounding is possible in such a cycle because all of the words have a
referent; however, because all references within such a closed cycle
appear to be exclusively kept within language and do not seem any
longer to refer to anything dynamic, there is a danger of there being
no meaning other than circularity to such a purely static construct.
Again, symbols are tools for modelling and are useful only if the
circularity is not mistaken for a grounded reference.  Thus, I avoid
strictly presenting such illusions of grounding as primitive in this
dissertation.  To allow for grounded descriptions, I will first
outline my conception of the dynamic, to which my description
ultimately must refer.  A description of reflective thinking referring
to an \emph{a~priori} concept that has not been derived in terms of
symbols allows for something to remain outside of language as the
dynamic referent.  This last step avoids the creation of a purely
self-referential cycle and prevents illusions from becoming mistaken
for grounded parts of the model.  In the next section I will describe
the necessary conception of the dynamic to which my description of
reflective thinking will refer.

\section{Activities in Duration}

\cite{bergson:1910}, in his seminal \underline{Time and Free Will},
discusses the dynamic as heterogeneous activities in Duration.
Activities in Duration will serve as my conception of the dynamic for
my description of reflective thinking.  Activities in Duration are the
dynamic ongoing activity, heterogeneous activities that are neither
distinct nor separate.  There is a trick, a sort of slight of hand,
required here in describing Bergson's conception of the activities in
Duration because these activities exist independently of
symbolization.

For example, when I look at a tree, I may say that the tree has green
leaves and a brown trunk; the symbols, green and brown, refer to the
activities in Duration, which I am seeing.  When I focus carefully on
the dynamic activity of my situation, I see that the leaves and trunk
could be described by a greater combination of various greens,
yellows, grays, browns and reds.  Symbols limit perception to a
discrete sort of projective presupposition.  Therefore, as I introduce
more symbolic references in my increasingly limited description of the
dynamic, a growing danger of losing focus on something dynamic
emerges, creating a cycle of symbols seemingly referring exclusively
to symbols.

The motivation to tangentially define more clearly what I mean by the
symbol green or brown by using more symbols results in a circularity
from which there is no exit, and therefore, a false dilemma from which
I will never again be able to see or talk about any dynamic tree as
the referent.  In order to avoid the danger of purely cyclic symbolic
references in my model, I will continue to relate the parts of my
model to the dynamic activity of reflective thinking, which exists as
activities in Duration.

\section{Consciousness, Awareness, and Experience}

There are many terms that are used in the literature when referring to
the concept of the dynamic.  Each of these terms has a different
associated collection of necessary absurdities.  For example, the term
consciousness begs the question \emph{of what} one is conscious.
There is an implied subject and object relationship when using this
transitive term, which is not meant when the term is used as an
isolated absolute reference to the dynamic.

The terms awareness and experience are used similarly and have the
same necessary absurdities as the term consciousness when they are
considered as absolutely singular references to the dynamic.  Each of
these terms has different meaningful uses when they are not used in
this absolutely singular sense.  For example, in this dissertation an
awareness of the potential confusions these terms introduce when they
are used singularly in abstract isolation as reference to the dynamic
will be maintained.

Thus, I will not argue that these conceptions of the dynamic are
incorrect, but in order to avoid confusion, I will never use these
terms in their abstracted sense.  Instead, I will conflate all of
these uses of these terms as all being the same as my conception of
the dynamic, the ongoing activity that my model accepts, symbolizes,
and thinks about as given.\footnote{Because my model of mind is an
  absolute reference to the dynamic, it makes very few modelling
  assumptions.  Because my model has been previously mistaken for the
  philosophy of ``Subjective Idealism'', I have included a description
  of the mistake of Subjective Idealism, and a clarification that my
  model does not make this mistake, an important difference between my
  model and this problematic philosophy; I explain in
  \autoref{section:the_mistake_of_subjective_idealism}.}

\section{Actual Reflection}

The activities in Duration are the given dynamic ongoing.  These
activities are the inseparable actions that are currently available to
reflection.  Reflection is the ability to symbolize this ongoing
activity.  For example, the activity of perceiving a table can be
symbolized as ``table''.  Similarly, the activity of walking, can be
symbolized as ``walking''.  Note that my model of reflective thinking
assumes that there is ongoing activity before any symbols exist for
the actions of thinking to manipulate.  Thinking is an auxiliary
activity that manipulates static symbolic constructions that refer to
the primary ongoing activity of existence.  Thus, the activity of
perceiving a table requires neither that this activity be reflectively
symbolized nor that a secondary thinking activity manipulates the
resulting symbols.  I will use the phrase \emph{actual reflection} to
refer to the activity of symbolizing ongoing activities.

\section{Order of Space}

Activities in Duration may be symbolized in a homogeneous medium
without activity.  Bergson calls this medium Space, allowing symbolic
references to activities in Duration to be related.  Of major
importance is the fact that one does not derive the other, but
activity and order coexist with neither Duration being a derivative of
Space nor Space containing the fundamental activities in Duration.
Obviously, Bergson has gone through painstaking efforts to leave room
in his description for the dynamic as some independent, non-contingent
activity not derived from either Space or Duration.  He makes the
necessary concession that a description of something dynamic is only
grounded in that it leaves room for this dynamic to exist
independently of its own symbolization.  This modelling work is meant
to be similarly grounded in the necessarily separate dynamic activity
of reflectively learning to accomplish goals.

\section{Modelling the Dynamic}
\label{section:modelling_the_dynamic}

Duration and Space are the fundamental duality that forms the basis of
my model of the dynamic activity of reflectively thinking.  Thus this
duality exists pre-reflectively, prior to symbols being created from
the activities in Duration and prior to these symbolic references
being related in Space.  Note the fundamental contradiction of
intentionally describing the dynamic only in static symbolic terms.
For example, if Space has no activity and is not itself composed of
activities in Duration, to what is the symbol ``Space'' referring?
Symbol ``Space'' refers only to the logical potential for an absence
of activity, the potential for itself as not existing; thus, it is the
logical necessity of the negative implication of positive existence,
to which no alternative can be assigned.  Again, my model will
explicitly maintain an awareness of this implicit, fundamental and
irresolvable contradiction.  Thus, my description will appear to begin
of something in the middle, the \emph{in medias res} of an underived
time and space, maintaining an explicit awareness that descriptions
are only tools for reflection.  Duration and Space are not
placeholders for the dynamic; they are useful, intermediate fictions
that I use here to model the dynamic ongoing, the activity of being,
that which is not a static arrangement of symbols.

\section{Spatial Reflection}

The activities in Duration are the dynamic activity of being in the
present, actively dynamic and ongoing.  Symbols are static references
to the dynamic present.  There is a fundamental distinction here
between the dynamic and the static.  In this way, Spatial arrangements
of symbols can actually exist in the dynamic as an actively maintained
static construction.  The ongoing activities that hold symbols in a
specific Spatial arrangement can be symbolically reified.  I will
refer to the activity of symbolizing the activities of a Spatial
arrangement of symbols as \emph{Spatial reflection} or
\emph{reification}.  Spatial reflection is a form of actual
reflection.

\section{The Physical Layer}

My model includes three layers of ongoing dynamic activity.  I will
use the phrase \emph{physical layer} to refer to the pre-reflective
layer of activities that exists necessarily prior to any subsequent
reflective thinking layers.  The physical layer is the only
pre-reflective layer of activity in my model.  Therefore, a clear line
is drawn between the physical layer and those activities that
symbolize and Spatially arrange symbols.  Thus, they are not included
in what I will refer to as physical activity.  Note that in most AI
models, there is a distinction between the mind and the body or the
environment.  In my model, \emph{all} activities that exist are part
of the mind.  The dynamic mental activities that I am referring to in
my model are not an object that is viewed subjectively.  Critical to
an understanding of the goals of objective science is that the
creation of object and subject distinctions is part of the activities
of the mind; because this understanding is so critical to an
advancement of scientific understanding, I will clarify the mental
derivation of objective science in \autoref{chapter:science}.  Because
the creation of object and subject relationships is a relatively
advanced mental activity and not key to my thesis, I will discuss this
as future research in \autoref{chapter:future}, which would be a
promising place to extend my model to include ideas like physical
bodies and other subjective and objective environmental perspectives.

\section{Orders of Reflective Layers}

In addition to the pre-reflective physical layer of activity, my model
includes two additional layers of reflective thinking activity that
use two different classes of causal models in order to accomplish two
different classes of goals.  I will use the phrase \emph{first-order
  reflective layer} to refer to the reflective layer of thinking that
creates symbols and Spatial arrangements that refer to physical
activities.  I will refer to reflectively symbolizing physical
activities as \emph{physical reflection}.  Physical reflection is a
form of actual reflection and is one of the primary reflective
thinking activities; remember that physical activities are
pre-reflective, so physical reflection is not itself a physical
activity.

Obviously, the first-order reflective layer cannot create symbols or
Spatial arrangements that refer to its own activities, with one
exception: Spatial reflection.  The first-order reflective layer has
the \emph{a~priori} ability to maintain its own actively reified
Spatial relationships as existing between symbols that refer to
physical activities.  The lack of the ability for the first-order
reflective layer to refer to its own symbolically creative or
Spatially manipulative activities clarifies my model by keeping clear
distinctions between each class of causal knowledge in each subsequent
reflective layer of thinking.  Before describing distinctions between
classes of causal models, I will spend the next few sections
describing the composition of temporal transitions and causal
hypotheses.

Reflection over the activity of the first-order reflective layer
necessarily creates a second layer of reflection and thinking.  Thus,
clearly deriving and describing this second layer of reflective
thinking additionally necessitates an implicit and arbitrary number of
layers.  This condition is the focus of this dissertation.
\cite{minsky:2006} describes a six-layer reflective model of thinking
about thinking.  The first three layers of my model and
implementation, including these two layers of reflective thinking, can
be compared to the first four layers of Minsky's model.  There are
many close parallels between my model and Minsky's model, and I will
describe these modelling parallels in
\autoref{chapter:related_models}.

\section{Extensive and Intensive Quantities}
\label{section:extensive_and_intensive_quantities}

\cite{bergson:1910} makes a key distinction between comparisons of
``extensive'' quantities and comparisons of ``intensive'' quantities.
Extensive quantities are Spatial comparisons.  Extensive quantities
are between two symbolic references that share a containment
relationship in terms of their referents: one symbolic referent is
inclusively contained by another symbolic referent in Space.  For
example, one body may be said to be larger than another in terms of
its Spatial extent.  Intensive quantities are more subtle and an
example is in terms of pain, which can be said to be more or less
intense.  Intensive quantities are similar to neither numerically nor
Spatially extensive containment relationships but, instead, refer to
different fundamental activities in Duration.  For example, a pain
that has no intensity is not a pain at all; a pain that has a mild
intensity may be equivalently referred to as an irritation or an itch;
a pain of high intensity may be equivalently referred to as a sharp
pain or a throbbing pain.  The point is that these intensities of
quality are actually very different from extensive quantities in that
they are not greater than or less than one another in the sense of a
containment relationship.  For example, a very intense pain, such as a
shooting pain or a throbbing pain, does not in any sense contain a
lesser mild pain, such as an irritation or an itch.  This same pattern
exists with detailed descriptions of other qualities that are often
thought of as having comparable intensities: heat and cold; light;
pressure; sound; pitch; aesthetic feelings, such as grace and beauty
in music, poetry and art; emotions, such as rage and fear; moral
feelings, such as pity; affective sensations, including pleasure, pain
and disgust.  Each of these examples points out that the apparent
relationships of greater than or less than with respect to intensity
are \emph{dynamically static Spatial arrangements} of different
fundamental activities in Duration.

\section{Number}
\label{section:number}

\cite{bergson:1910} shows the derivation of all numerical
representations as requiring an accompanying extension in Space.  For
example, in referring to the individual sheep in a flock of sheep, he
explains counting:

% pg. 75-77
\begin{quote}
Number may be defined in general as a collection of units, or,
speaking more exactly, as the synthesis of the one and the
many\ldots

No doubt we can count the sheep in a flock and say that there are
fifty, although they are all different from one another and are easily
recognized by the shepherd: but the reason is that we agree in that
case to neglect their individual differences and to take into account
only what they have in common.  On the other hand, as soon as we fix
our attention on the particular features of objects or individuals, we
can of course make an enumeration of them, but not a total\ldots

Hence we may conclude that the idea of number implies the simple
intuition of a multiplicity of parts or units, which are absolutely
alike.

And yet they must be somehow distinct from one another, since
otherwise they would merge into a single unit.  Let us assume that all
the sheep in the flock are identical; they differ at least by the
position which they occupy in space, otherwise they would not form a
flock.
\end{quote}

The last point here illustrates the idea that numbers are always
derived from symbols situated in Space.  Therefore, a model for
thinking based directly on the symbols required cannot be more
complicated than a model based on a numerical representation.  A
number is, after all, the reified constructive result of an assumed
Spatial organization of symbolic references to the dynamic referent,
specific dynamic sheep.  The explicit implication is, therefore, that,
in building an AI that is capable of reflective thought, model
building and the use of numerical representations must be explicitly
available for inspection by the AI itself.  Here, I employ Occam's
razor to simplify the loop of reflective representation in order to
ease building the first proof-of-concept examples of AIs capable of
reflective thinking.

\section{Time as Space}

Symbolic references can be ordered Spatially.  These symbolic
relationships can be reified and treated symbolically by further
ordering Spatial relationships.  Thus, Bergson explains time as a form
of Space:

% pg. 98
\begin{quote}
Now, if space is to be defined as the homogeneous, it seems that
inversely every homogeneous and unbounded medium will be space.  For,
homogeneity here consisting in the absence of every activity, it is
hard to see how two forms of the homogeneous could be distinguished
from one another.  Nevertheless it is generally agreed to regard time
as an unbounded medium, different from space but homogeneous like the
latter: the homogeneous is thus supposed to take two forms, according
as its contents co-exist or follow one another.  It is true that, when
we make time a homogeneous medium in which conscious states unfold
themselves, we take it to be given all at once, which amounts to
saying that we abstract it from duration.  This simple consideration
ought to warn us that we are thus unwittingly falling back upon space,
and really giving up time.
\end{quote}

Time exists as a static creation, a derivative arrangement of symbols
in Space.  Understanding time to be static is important because this
allows the dynamic heterogeneous activities in Duration and a static
homogeneous unqualified Space to remain, respectfully, neither
distinct nor separate.  By prohibiting this separation, and, thereby,
their independent existences, I avoid defining the dynamic activities
in Duration in terms of static symbols, avoiding the creation of a
self-referential cycle, thus sidestepping the potential illusion of
dynamism by being clear in our understanding of time used
intentionally as a static tool of thought.

In my model, symbols are put into a static Spatial arrangement that I
refer to as time.  The actions that create symbols and put these
symbols into Space are basic parts of my model of reflective thinking.

\section{Temporal Reflection}

Spatial arrangements of symbols can be used to represent a past
through memory and subsequently project an analogous ``future'',
through imaginative action, thus constructing a derivative conception
of time.  The past and the future, therefore, remain static
constructions that are actively maintained as Spatial relationships in
the present.  I will refer to the reification of Spatial relationships
that order the past and the future as \emph{temporal reflection}.
Sequential time is created through temporal reflection.  Temporal
reflection is a form of Spatial reflection.  Like Spatial reflection,
temporal reflection is also a form of actual reflection, symbolizing
the active maintenance of a temporal Spatial relationship between
static symbolic references.

\section{Simultaneity and Transition}

Correlation requires two symbols to be related in Space.  As symbols
are correlated in the Space called time, these temporal correlations
are referred to as either simultaneities or transitions.
Simultaneities represent two symbols that are both actively symbolized
in Duration.  Transitions represent the change from the past to the
future.  Transitions can be extended in time in order to create
counterfactual temporal extensions, called inferences.  Inferences of
the past and the future can be created by matching and repeating these
transitions.  A transition implies the removal of one symbol and the
addition of another as a progression is made through a temporal
sequence.  If correlations are counted and compared in ratios, then
these are called probabilistic correlations, and the resulting
inference would be a probabilistic inference.

