A system built on a layered reflective cognitive architecture presents
many novel and difficult software engineering problems.  Some of these
problems can be ameliorated by erecting the system on a substrate that
implicitly supports tracing of results and behavior of the system to
the data and through the procedures that produced those results and
that behavior.  Good traces make the system accountable; it enables
the analysis of success and failure, and thus enhances the ability to
learn from mistakes.

This constructed substrate provides for general parallelism and
concurrency, while supporting the automatic collection of audit trails
for all processes, including the processes that analyze audit trails.
My system natively supports a Lisp-like language.  In such a language,
as in machine language, a program is data that can be easily
manipulated by a program, making it easier for a user or an automatic
procedure to read, edit, and write programs as they are debugged.

Constructed within this substrate is an implementation of the bottom
four layers of an Emotion Machine cognitive architecture, including
built-in reactive, learned reactive, deliberative, and reflective
layers.  A simple natural language planning language is presented for
the deliberative control of a problem domain.  Also, a number of
deliberative planning algorithms are implemented in this natural
planning language, allowing a recursive application of reflectively
planned control.  This recursion is demonstrated in a fifth
super-reflective layer of planned control of the reflective planning
layer, implying $N$ reflective layers of planned control.

Here, I build and demonstrate an example of reflective problem solving
through the use of English plans in a block building problem domain.
In my demonstration an AI model can learn from experience of success
or failure.  The AI not only learns about physical activities but also
reflectively learns about thinking activities, refining and learning
the utility of built-in knowledge.  Procedurally traced memory can be
used to assign credit to those thinking processes that are responsible
for the failure, facilitating learning how to better plan for these
types of problems in the future.
