%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

%Don't do anything that isn't play. \\ \medskip
%    --- Joseph Campbell



%\begin{flushright}{\slshape    
%Don't do anything that isn't play.} \\ \medskip
%    --- Joseph Campbell
%\end{flushright}

\begin{flushright}{\slshape    
Eternity has nothing to do with time. Eternity is that dimension \\
of here and now which thinking and time cuts out. This is it. \\
And if you don't get it here, you won't get it anywhere.} \\ \medskip
    --- Campbell (\citeyear{campbell:1988})
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

Push Singh for being a strong memory of a heroic friend and advisor.

\vspace{5mm}

\noindent Joseph Paradiso for his advice, encouragement and support at all times.

\vspace{5mm}

\noindent My readers for both being inspirations as well as tirelessly correcting countless drafts:
Gerry Sussman,
Michael Cox, and
Marvin Minsky.

\vspace{5mm}

\noindent My immediate family:
Greg Morgan, Carolyn Spinner, Paul Bergman, Virginia Barasch, and Leaf Morgan.

\vspace{5mm}

\noindent Those who worked on the project:
Dustin Smith for the physical social commonsense kitchen simulation.
Radu Raduta for the PID control loops for robot balance and movement.
Jon Spaulding for the EEG amplifiers and pasteless electrodes.
Gleb Kuznetsov for the first version of a 3D commonsense kitchen critic-selector layout planner.
Jing Jian for being so good at finding critical bugs in Funk2 and for writing all of the learned-reactive resources for following recipes in the kitchen.
Panupong Pasupat for the error-correcting subgraph isomorphism algorithm and the Winston difference-of-difference analogy algorithm.
Mika Braginsky for the visualization of mental resources, agencies, and layers.

\vspace{5mm}

\noindent For their encouragement, advice, and inspiration:
Walter Bender,
Henry Lieberman,
Patrick Winston,
Whitman Richards,
Aaron Sloman,
Ed Boyden,
Sebastian Seung,
Ted Selker,
Hugh Herr, and
Rebecca Saxe.

\vspace{5mm}

\noindent The Responsive Environments Group:
Joe Paradiso,
Amna Cavalic,
Josh Lifton, % for inspiring me to write my first garbage collector and ultimately my own programming language.
Mat Laibowitz, % for disagreeing with me about almost everything, being the hardest-core engineer, and refusing ``to party''.
Mark Feldmeier, % for being there to ask questions, listen, and talk.
Jason LaPenta,
Matt Aldrich, % for jamming, teaching me about metal, learning technically intricate music.
Nan-Wei Gong, % for singing, playing keys, and generally rocking out.
Mike Lapinski, % for being a healthy athlete.
Gershon Dublon,
and Nan Zhao. % for dancing and playing music.

\vspace{5mm}

\noindent The Common Sense Computing Group:
Push Singh,
Henry Leiberman,
Walter Bender,
Marvin Minsky,
Felice Gardner,
Betty Lou McClanahan,
Catherine Havasi,
Rob Speer,
Hugo Liu,
Barbara Barry,
Ian Eslick,
Jason Alonso, and
Dustin Smith.

\vspace{5mm}

\noindent The Mind Machine Project:
Newton Howard, Gerry Sussman, Patrick Winston, Marvin Minsky, Neil Gershenfeld, Forrest Green, Kenneth Arnold, Dustin Smith, Catherine Havasi, Scott Greenwald, Rob Speer, Peter Schmidt-Nielsen, Sue Felshin, David Dalrymple, and Jason Alonso.

\vspace{5mm}

\noindent The Popes of GOAM:
Dustin Smith,
Scotty Vercoe,
Dane Scalise, and
Kemp Harris.

\vspace{5mm}

\noindent My adoptive thesis finishing family:
Lindsey and baby Sudbury, Andrew Sudbury, Jon Sudbury, and Todd Rautenberg.

\vspace{5mm}

\noindent
The following friends for laughing with me over the long haul:
Dustin Smith, % for being a rock.
Grant Kristofek, % for loving music, staying true to the blues, and keeping my soul alive by organizing jams.
Rhett Nichols,
Mat Laibowitz, % for disagreeing with me about almost everything, being the hardest-core engineer, and refusing ``to party''.
Nan-Wei Gong, % for singing, playing keys, and generally rocking out.
Scotty Vercoe,
Simon LaFlame,
Dane Scalise,
Hannah Perner-Wilson, % for discussing the limitations of categorical thinking.
Clio Andris, % for going through the sometimes painful PhD process with me.
Nick Dufour,
Jeff Lieberman, % for practicing simply being.
Mark Feldmeier, % for being there to ask questions, listen, and talk.
Matt Aldrich, % for jamming, teaching me about metal, learning technically intricate music.
Edwina Portocarrero, % for being deep.
Mike Lapinski, % for being a healthy athlete.
Nan Zhao, % for dancing and playing music.
Mako Hill, % for helping me get Funk2 out into the public open software development communities.
David Cranor, and % for organizing thesis crew.
Em\"{o}ke-\'{A}gnes Horv\'{a}t. % for being magical.

\vspace{5mm}

\vspace{5mm}

\noindent
This research was supported by an allocation of advanced computing
resources provided by the National Science Foundation.  Some of the
computations were performed on Nautilus at the National Institute for
Computational Sciences (\url{http://www.nics.tennessee.edu/}).

\endgroup

